### 2012-12-28 edits/barryedits.tex
Added some more of Barry's edits.

\documentclass[a4paper,12pt]{article}

\usepackage{../mathbf-abbrevs}
\input{../defs}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

\usepackage[square,comma,numbers,sort&compress]{natbib}

\begin{document}

\begin{enumerate}
\item Page 1. Column 2. I've left the sentence: ``The idea is to `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.'' in.  If I remove it, Bill and Andre will want to put it back.  I believe it does give the correct intuition, even if it is imprecise.
%\item Regarding the ``conditioned upon minimisation with respect to'' phrases, I have replaced ``conditioned upon'' with ``concentrated by''.  This is perhaps a bit lazy, but I 
\item Page 6, column 2, you want me to add reference for the polynomial phase estimation papers, where we talk about differentiability of the objective function at the estimator.  I've added a reference to the article on arxiv.  We can change it to the trans. sig. proc. papers later.
\item Page 6, column 2, have left in the definition of $o_P(1)$, since it doesn't hurt.
\item Page 10 column 2, I have left ``be $N$ points uniformly spaced on the interval $[-\pi, \pi)$'' since it doesn't hurt.
\item Page 11, column 1, about the phrase ``where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$''.  I want to leave this.  The phrase is taken from one of David Pollard's papers.  It is informal, but I know that if I were personally reading this for the first time, I would be scratching my head at this point until I realised that expectation has been moved outside the supremum.
\item Page 11, column 2, about the Lipschitz continuous statement.  Obviously the term involving $\cos(\phi)$ needs only the mean value theorem.  However the term involving $\cos\fracpart{\phi}$ isn't always differentiable so the mean value theorem doesn't always apply.  It's obviouly Lipschitz continuous though, just plot it.  We can add a proof if you really think it's necessary.
\item Page 12, column 1, ``for $i \in D$ such that $R_i > 0$.'' I think this is necesary, otherwise the surrounding statement is not always true.
\item Lemma 11, I would also like there to be a nicer proof!  I haven't found one though.
\item 
\end{enumerate}

\end{document}