### 2014-07-03 responses/proofcorrection.tex
added more changes.  Past the sectioning error on page 7 now.

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs}

\usepackage{zref-xr,zref-user}
\zexternaldocument*{paper}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, I.\ Vaughan L.\ Clarkson and Barry Quinn  
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

%\maketitle

\begin{enumerate}
\item Page 1, column 2.  The formating of the first equation is wrong.  Space between $p_i$ and $i$ and $d_i$ and $i$ must be included.  It should look like:
\[
s_i = \begin{cases}
p_i & i \in P \\
d_i & i \in D,
\end{cases}
\]

% \item Page 4, bottom of column 1.  The equation beginning $Y_1 = B + \sum_{i \in D} g_{1i}$ is broken up.  This need to all be in one, i.e., like:
% \begin{align*}
% Y_1 &= B + \sum_{i \in D} g_{1i} \\
% &= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
% &= Y_0 + \eta g_{0\sigma(0)},
% \end{align*}

\item Page 4, column 2.  The phrase ``the loops on lines 1 and II'' the II should be 11 (the number eleven rather than two capital I's).  That is, ``the loops on lines 1 and 11''.

\item Page 5, column 1.  The phrase ``to denote $x$ taken `~modulo $2\pi$' into the interval \dots''.  There is a space between the apostrophe and the word modulo.  That is `~modulo $2\pi$' shoud be `modulo $2\pi$'.

\item Page 6, column 1, start of section 4.  The first equation starting $LS(\rho, \theta) = \dots$ has the word `where' placed in the middle of the column.  The word `where' should be left justified rather than centered.  It should look like:
 \begin{align*}
LS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
\end{align*}
where
\[
Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
\]

\item Page 6 column 2 second last equation starting $Z(\theta) = \dots$.  The equation ends with a full stop.  The full stop is put into the superscript of the equations, but it should not be.  It should look like:
\[
Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}.
\]

\item On page 7 the section entitled ``Proof of asymptotic normality (Theorem~2)'' does not have a section number.  It should be V (i.e. the roman numeral for 5).  The sections that follow entitled ``Simulations'' and ``Conclusion'' and correspondingly numbered incorrecty.  The ``Simulations'' section should be section VI (i.e. 6) and the ``Conclusion'' section should be VII (i.e. 7).

\item

\end{enumerate}

\end{document}