### 2014-07-04 responses/proofcorrection.tex
minor

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs}

\usepackage{zref-xr,zref-user}
\zexternaldocument*{paper}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, I.\ Vaughan L.\ Clarkson and Barry Quinn  
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

%\maketitle

\begin{enumerate}
\item Page 1, column 2.  The formatting of the first equation is wrong.  Space between $p_i$ and $i$, and $d_i$ and $i$ need to be included.  It should look like:
\[
s_i = \begin{cases}
p_i & i \in P \\
d_i & i \in D,
\end{cases}
\]

% \item Page 4, bottom of column 1.  The equation beginning $Y_1 = B + \sum_{i \in D} g_{1i}$ is broken up.  This need to all be in one, i.e., like:
% \begin{align*}
% Y_1 &= B + \sum_{i \in D} g_{1i} \\
% &= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
% &= Y_0 + \eta g_{0\sigma(0)},
% \end{align*}

\item Page 4, column 2.  The phrase ``the loops on lines 1 and II'' the II should be 11 (the number eleven rather than two capital I's).  That is, ``the loops on lines 1 and 11''.

\item Page 5, column 1.  The phrase ``to denote $x$ taken `~modulo $2\pi$' into the interval \dots''.  There is a space between the apostrophe and the word modulo.  That is `~modulo $2\pi$' should be `modulo $2\pi$'.

\item \label{com:leftjustwhere} Page 6, column 1, start of section 4.  The first equation starting $LS(\rho, \theta) = \dots$ has the word `where' placed in the middle of the column.  The word `where' should be left justified rather than centered.  It should look like:
 \begin{align*}
LS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
\end{align*}
where
\[
Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
\]

\item Page 6 column 2 second last equation starting $Z(\theta) = \dots$.  The equation ends with a full stop.  The full stop is put into the superscript of the equation, but it should not be.  It should look like:
\[
Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}.
\]

\item On page 7 the section entitled ``Proof of asymptotic normality (Theorem~2)'' does not have a section number.  It should be V (i.e. the roman numeral for 5).  The sections that follow entitled ``Simulations'' and ``Conclusion'' are correspondingly numbered incorrectly.  The ``Simulations'' section should be section VI (i.e. 6) and the ``Conclusion'' section should be VII (i.e. 7).

\item Page 10, bottom of first column, equation beginning $LS(\{r_i, i \in D\})$.  This equation would preferably not be split over multiple columns, i.e., it should all be in a single column like:
\begin{align*}
LS&(\{r_i, i \in D\}) \\
&= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
&= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
&= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
&= LS(\{\hat{d}_i, i \in D\}) - C,
\end{align*}
% If you can make it fit then:
% \begin{align*} 
% LS(\{r_i, i \in D\}) &= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
% &= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
% &= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
% &= LS(\{\hat{d}_i, i \in D\}) - C,
% \end{align*}
% would be even better.

\item Page 10 column 2, between first and second equation.  The word `where' is centered rather than left justified.  This is similar to my comment~\ref{com:leftjustwhere}.

\item Page 10 column 2, half way down.  The word `and' is centered rather than left justified.  This is similar to the above comment and my comment~\ref{com:leftjustwhere}.  It should look like:
\begin{align*}
\eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}} &= (e^{-j2\pi/M} - 1) e^{j \pi/M} e^{jv} \\
&= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
&= -2 j \sin(\tfrac{\pi}{M}) e^{j v },
\end{align*}
and
\[
\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) = 2 \sin(\tfrac{\pi}{M}) \sin(v).
\]

\item Page 11, top of column 2, between first and second equations.  The word `where' is centered when it should be left justified.

\end{enumerate}

\end{document}